---
title: "Tidybiology +DS: Session 1 "
subtitle: "An Introduction to Biological Data Science in R"
author: "Matthew Hirschey, Ph.D."
date: "April 8-9, 2020"
output: 
  ioslides_presentation:
    widescreen: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(knitr)
library(here)
library(twitterwidget)
library(rlang)
knitr::opts_chunk$set(echo = FALSE, 
                      fig.align="center")
```

```{r load_data, message=FALSE, warning=FALSE, include=FALSE}
#df input
file_name <- "heart.Rds" #set this 
file_label <- "Heart Disease Dataset" #set this 
dataframe_name <- "heart" #set this 
path_to_data <- here::here("data", file_name)
data <- readRDS(path_to_data)
heart <- data #set this name, makes it available in the learnr chunks
#heart_joined
#patient_hx

#set.seed(42)
df_input <- data # df MUST contain at least one character and one numeric variable
#df_input <- df_input[ , sample(ncol(df_input))] # shuffle columns 
df_numeric <- df_input %>% select_if(is.numeric)
df_character <- df_input %>% select_if(is.character)
df_numeric_select <- df_numeric[sample(1:ncol(df_numeric), 1)] # choose a random numeric column
df_numeric_filter <- df_numeric_select[sample(1:nrow(df_numeric), 1),] # choose a random number within that column
df_character_select <- df_character[sample(1:ncol(df_character), 1)] # choose a random character column
df_character_filter <- df_character_select[sample(1:nrow(df_character), 1),] # choose a random value within that column
```

## Watson Health/AI parable
```{r intro, echo=FALSE, fig.cap="", out.height = '100%'}
url <- "https://source.unsplash.com/701-FJcjLAQ/500x500"
knitr::include_graphics(url)
```
<font size = "1">Photo by [National Cancer Institute on Unsplash](https://unsplash.com/@nci)</font>

## commodity is new knowledge

## how do you come up with a new idea?

## scientific methods
- something about RCTs

## Hierarchy of Knowledge Gathering
```{r hierarchy, echo=FALSE, fig.cap="", out.height = '100%'}
url <- "https://upload.wikimedia.org/wikipedia/commons/0/06/DIKW_Pyramid.svg"
knitr::include_graphics(url)
```
<font size = "1">https://en.wikipedia.org/wiki/DIKW_pyramid</font>

## Data Science Tools

## Emerging field of Data Science
```{r venn1, echo=FALSE, fig.cap="", out.height= '100%'}
url <- "https://images.squarespace-cdn.com/content/v1/5150aec6e4b0e340ec52710a/1364352051365-HZAS3CLBF7ABLE3F5OBY/ke17ZwdGBToddI8pDm48kB2M2-8_3EzuSSXvzQBRsa1Zw-zPPgdn4jUwVcJE1ZvWQUxwkmyExglNqGp0IvTJZUJFbgE-7XRK3dMEBRBhUpxPe_8B-x4gq2tfVez1FwLYYZXud0o-3jV-FAs7tmkMHY-a7GzQZKbHRGZboWC-fOc/Data_Science_VD.png?format=1500w"
knitr::include_graphics(url)
```
<font size = "1">http://drewconway.com/zia/2013/3/26/the-data-science-venn-diagram</font>

## Venn Diagram of Data Science v2.0
```{r venn2, echo=FALSE, fig.cap="", out.height= '100%'}
url <- "https://3.bp.blogspot.com/-bvQxcwfqATQ/V-E_uTBc4VI/AAAAAAAAMGQ/Qa1Ntef-rs0E-mWx5pkVu-CPlREdvD0TwCLcB/s1600/VennDiagram2.png"
knitr::include_graphics(url)
```
<font size = "1">Joel Grus via KDnuggets</font>

## Overall goal is Knowledge Generation
```{r knowledge, echo=FALSE, fig.cap="", out.height = '100%'}
url <- "https://upload.wikimedia.org/wikipedia/commons/0/06/DIKW_Pyramid.svg"
knitr::include_graphics(url)
```
<font size = "1">https://en.wikipedia.org/wiki/DIKW_pyramid</font>

## Several Approaches to Knowledge Generation
```{r approaches, echo=FALSE}
tweet <- twitterwidget('1125268670324695041')
```
<center>`r tweet`</center>

## World's most popular programming languages
```{r popular, echo = FALSE}
excel <- tibble(
  name = c("Excel", "Java", "C", "C++", "Python"), 
  num = c(100000000, 9000000, 6000000, 4000000, 3000000)
)
excel_plot <- ggplot(excel) +
  geom_col(aes(x = fct_rev(fct_reorder(name, num)), y = num), fill = "navy") +
  labs(x = "", y = "Number of Users (Million)") +
  scale_y_continuous(labels = c("0", "25", "50", "75", "100")) +
  theme_minimal()
excel_plot
#need to add source
```

# R <br> language

## R is a language
```{r language, echo=FALSE, fig.cap=""}
url <- "https://source.unsplash.com/ZzWsHbu2y80/256x455"
knitr::include_graphics(url)
```
<font size = "1">Photo by Hannah Wright on Unsplash</font>

## R has values
- 1
- "North Carolina"
- "2020-04-08"

## R has objects
- A name without quotes  
- Assigned using `<-` (looks like an arrow pointing left)  
- Can be a value, object, or function result  

## Try assigning an object 
1. Assign an object; **remember, no quotes on name**  
`name <- 4`  
  
2. Return that object by typing its name  
`name`  
  
Try this in the code chunk below, then hit "Run Code"  

```{r my_first_object, exercise=TRUE}

```

## R has functions
- A name without quotes  
- Followed by () to run the function  
- Optional arguements: values, objects, or function results  
- `round(x, digits = 3)`

Try this in the code chunk below, then hit "Run Code"  

```{r my_first_fun, exercise=TRUE}
round(pi, digits = 3)

```

## Which one of these are numbers?

```{r num_quiz, echo=FALSE}
question("Which of these are numbers?",
  answer("1", correct = TRUE),
  answer('"1"', message = "Because it has quotes, it is an string"),
  answer('"one"', message = "Because it has quotes, it is an string"),
  answer("one",  message = "R recognizes this as an object"), 
  allow_retry = TRUE, 
  random_answer_order = TRUE
)
```

## Which of these will work?
Suppose `one <- 1`
```{r one_quiz, echo=FALSE}
question("Which of these are numbers?",
  answer("log(1)", correct = TRUE),
  answer('log("1")', message = "You cannot log transform a word (string)"),
  answer('log("one")', message = "You cannot log transform a word (string)"),
  answer("log(one)", correct = TRUE), 
  allow_retry = TRUE, 
  random_answer_order = TRUE
)
```
Try it for yourself!  
  
```{r scratchpad, exercise=TRUE}
one <- 1
```

## Data are stored in tables and dataframes
Data stored in a dataframe are conceptually equivalent to a spreadsheet with rows and columns  
  
This is a sample from the ``r dataframe_name`` dataset  
  
```{r kable}
df_input %>% 
  select(1:5) %>% 
  slice(1:5) %>% 
  knitr::kable()
```

## Data are stored in tables and dataframes
Data stored in a dataframe are conceptually equivalent to a spreadsheet with rows and columns  
  
This is a sample from the ``r dataframe_name`` dataset  
  
```{r tibble}
df_input %>% 
  select(1:5) %>% 
  slice(1:5) %>% 
  print(as_tibble())
```

## Extract or create new objects
```{r extract_setup, include=FALSE}
col <- sample(colnames(df_input), 1)
```

You can call a single part of the data frame  
  
`r dataframe_name`$`r col`

```{r extract, echo = FALSE, warning = FALSE}
df_input %>% 
  select(!!col)
```

## Extract or create new objects 
  
Write the R code required to extract a variable from the `r dataframe_name` dataset:  
  
Remember, the format is: `r dataframe_name`$`r col`  
  
```{r extract_exercise, exercise=TRUE}

```

## Extract or create new objects 
  
You can also save a part of the dataframe as an object for later use  
  
``r col` <- `r dataframe_name`$`r col``  
  
In the code chunk below:  
1. On the first line, write the R code to save a single column to a new object   
2. On the second line, type the object name #this will print out the new object  
3. Run the code  
  
```{r create_exercise, exercise=TRUE}

```

# R <br>Integrated Developer Environment

## R
Statistical Programming Language
```{r r_exercise, exercise=TRUE}

```
  
```{r r_image, echo=FALSE, fig.cap="", out.height = '100%'}
url <- "https://source.unsplash.com/Vb8pR6cKCRI/300×200"
knitr::include_graphics(url)
```
<font size = "1">Photo by [Photo by Evan Hein on Unsplash](https://unsplash.com/@evanheinphoto)</font>

## Rstudio
Integrated Developer Environment

```{r r_de, echo=FALSE, fig.cap="", out.height = '100%'}
url <- "https://source.unsplash.com/XuDPnpox8tc/300×200"
knitr::include_graphics(url)
```
<font size = "1">[Photo by Jaromír Kavan on Unsplash](https://unsplash.com/@jerrykavan)</font>

---
[Open Rstudio](http://www.rstudio.cloud)  
<iframe src="http://www.rstudio.cloud"></iframe>

## Rstudio Demonstration
Go to code/  
Open 01_r_demo.Rmd  
Follow along.  

# R <br>Markdown


## R Markdown Introduction
Go to code/  
Open 02_rmd_exercise.Rmd  
Read through the file and do everything it tells you to do.  

## R Markdown
An authoring format for Data Science.  
```{r markdown, echo=FALSE, fig.cap="", out.height = '100%'}
url <- "https://www.dropbox.com/s/ezsb12huhsx6iks/setup.jpeg"
knitr::include_graphics(url)
```

# R <br>Packages

## R Packages
The R language contains thousands of functions, data sets, and help pages.  
- but only a few hundred are included when you download R  
  
**This is called 'Base R'**  
  
The other functions, data sets, and help pages are grouped into collections known as packages that you can choose to download or not download.  

## “Verbs” (i.e. functions) act on data

`do_this(to_that)`  
  
`do_this(to_that, using_these)`  
   
*We talked about functions before (e.g. `round(pi, 3)`)*  
  
  
  
  
**Functions are the power of using R**

## Packages contain functions, documentation, data
  
<center><iframe src="https://dplyr.tidyverse.org"></iframe></center>

## CRAN
- Most R packages are stored on CRAN, alongside R.  
- Think of them as optional extensions of the R language.  
  
```{r cran, echo=FALSE, fig.cap="", out.height = '66%', out.width= '66%'}
url <- "https://gist.githubusercontent.com/daroczig/3cf06d6db4be2bbe3368/raw/b66b0531fb1b86d3e04a003b2e105ad4f147900e/number-of-submitted-packages-to-CRAN.png"
knitr::include_graphics(url)
```
<font size = "1">Image by [daroczig](https://gist.github.com/daroczig/3cf06d6db4be2bbe3368)</font>
  
## Using Packages (Part I)
  
1. `install.packages("tidyverse")`  
  
Do this 1 time per computer.  
  
This command will install the package into your instance of R, whether it is local, on a server, or in the cloud. This is required to use the functions in a package.  
  
## Tidyverse
  
<center><iframe src="https://www.tidyverse.org"></iframe></center>

## Tidyverse Pop Quiz!

The tidyverse contains the following packages (ggplot2, dplyr, tidyr, readr, purrr, tibble, hms, stringr, lubridate, forcats, DBI, haven, httr, jsonlite, readxl, rvest, xml2, modelr, tidyverse).   
  
  
  
  
**How would you  install them?**

## Tidyverse Pop Quiz!
`install.packages("ggplot2")`
`install.packages("dplyr")`
`install.packages("tidyr")`
`install.packages("readr")`
`install.packages("purrr")`
`install.packages("tibble")`
`install.packages("hms")`
`install.packages("stringr")`
`install.packages("lubridate")`
`install.packages("forcats")`
`install.packages("DBI")`
`install.packages("haven")`
`install.packages("httr")`
`install.packages("jsonlite")`
`install.packages("readxl")`
`install.packages("rvest")`
`install.packages("xml2")`
`install.packages("modelr")`
`install.packages("broom")`  
  
  
  
Better:  
`install.packages("tidyverse")`  
An R package that serves as a short cut for installing and loading the components of the tidyverse.

## Using Packages (Part II)
  
1. `install.packages("tidyverse")`  
  
Do this 1 time per *computer*.  
  
2. `library(tidyverse)`  
  
Do this 1 time per *session*  
  
## Using Packages (Part II)
Downloading a package isn't the same as using it. If you'd like to use an R package, you need to tell R. You do that by running the command library, again followed by parentheses and the package name. library loads all of the functions, data sets and help pages that come with the package into your R session, where you can use them. If you close R, you'll need to reload the package with library() if you want to use it again.  
  
  
## Using Packages (Part III)
# Setup
The setup chunk is always run once before anything else  



## next slide